---
title: "Data Visualisation"
subtitle: "RAdelaide 2016"
author: "Steve Pederson"
date: "20 July 2016"
output: 
  ioslides_presentation:
    css: ../custom.css
    fig_caption: yes
    keep_md: yes
    logo: ../images/UoA_logo_col_vert.png
---

```{r, loadPackages, echo = FALSE, include = FALSE}
library(knitr)
library(readr)
library(dplyr)
library(stringr)
library(reshape2)
opts_chunk$set(echo = TRUE, include = TRUE, 
               warning = FALSE, message = FALSE, fig.align = "center",
               results = 'hide', fig.show = 'hide')
data <- read_csv("data/comments.csv", comment = "#")
pcr <- read_csv("data/PCR.csv")
pcr_long <- pcr %>%
  melt(id.vars = "Gene",
       variable.name = "CellType", value.name = "Ct") %>%
  mutate(TimePoint = str_extract(CellType, "(0hr|12hr|24hr)"),
         CellType = str_replace(CellType, "_(0hr|12hr|24hr)", "")) 
```

# Data Visualisation | Using the Package `ggplot2`

## The package `ggplot2`

`R` has numerous plotting functions in the base package `graphics`

```{r, eval = FALSE}
?plot
?boxplot
?hist
```

__Go to the bottom of each help page and copy a few lines__

## The package `ggplot2`

- `ggplot2` gives much more flexibility and power
- Has unique syntax and approach
- We add layers of plotting information

```{r}
library(ggplot2)
```

## The package `ggplot2` | aesthetics

The main function is `ggplot()`

- In this first stage we set the plotting aesthetics using `aes()`
- This amounts to: What goes on the x & y axes, what defines the colour/shape etc.

```{r, fig.show='hide'}
ggplot(data, aes(x = weight, y = height))
```

When we do this, nothing will be plotted. We get the plot area only...

## The package `ggplot2` | geometry

After defining the plot aesthetics, we add the geometry using various `geom_...()` functions

We tell the first line that "more is to come" by adding a `+` symbol at the end

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point()
```

## The package `ggplot2` | geometry

```{r, echo=FALSE, fig.width=4, fig.height=4, fig.show='asis'}
ggplot(data, aes(x = weight, y = height)) +
  geom_point()
```

## The package `ggplot2` | aesthetics

There are numerous aesthetics available for `geom_point()`

```{r, eval=FALSE}
?geom_point
```

```{r, fig.show='hide'}
ggplot(data, aes(x = weight, y = height, colour = transport)) +
  geom_point()
```

```{r}
ggplot(data, aes(x = weight, y = height, 
                 colour = transport, shape = gender)) +
  geom_point()
```

## The package `ggplot2` | aesthetics

It can be a good idea to put the general aesthetics in `ggplot()`, with the `geom_point()` specific ones in that line

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point(aes(colour = transport, shape = gender))
```

Point aesthetics set outside of `aes()` are general across all the data

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point(aes(colour = transport, shape = gender), size = 4)
```

## The package `ggplot2` | adding multiple geoms

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point(aes(colour = transport, shape = gender)) +
  geom_smooth()
```

This defaults to a `loess` fit

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point(aes(colour = transport, shape = gender)) +
  geom_smooth(method = "lm", formula = y~x, se = FALSE)
```

## The package `ggplot2` | labels

Axis and legend labels can be added using `labs()`

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point(aes(colour = transport, shape = gender)) +
  geom_smooth(method = "lm", formula = y~x, se = FALSE) +
  labs(x = "Weight (kg)", y = "Height (cm)", 
       shape = "Gender", colour = "Transport")
```


## The package `ggplot2` | facets

(This is my favourite feature)

```{r}
ggplot(data, aes(x = weight, y = height)) +
  geom_point(aes(colour = transport, shape = gender)) +
  geom_smooth(method = "lm", formula = y~x, se = FALSE) +
  labs(x = "Weight (kg)", y = "Height (cm)", 
       shape = "Gender", colour = "Transport") +
  facet_wrap(~gender) 
```

## The package `ggplot2` | Different geoms

Enter `geom_` in the Console followed by the `tab` key

```{r}
ggplot(data, aes(x = height, fill = gender)) +
  geom_density(alpha = 0.5)
```

```{r}
ggplot(data, aes(x = gender, y =height, fill = gender)) +
  geom_boxplot()
```

## The package `ggplot2` | Different geoms

We could extract summaries from our data, then _pipe_ them into `ggplot()`

```{r}
data %>%
  group_by(transport, gender) %>%
  summarise(mn_height = mean(height), sd_height = sd(height)) %>%
  ggplot(aes(x = transport, y = mn_height, fill = transport)) +
  geom_bar(stat = "identity") +
  facet_wrap(~gender) +
  guides(fill =FALSE)
```

`geom_bar()` requires `stat = "identity"`

## The package `ggplot2` | Different geoms

We can add error bars

```{r}
data %>%
  group_by(transport, gender) %>%
  summarise(mn_height = mean(height), sd_height = sd(height)) %>%
  ggplot(aes(x = transport, y = mn_height, fill = transport)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mn_height - sd_height,
                    ymax = mn_height + sd_height),
                width = 0.6)+
  facet_wrap(~gender) +
  guides(fill =FALSE)
```

## The package `ggplot2` | Different geoms

Let's move the PCR data using `pcr_long`

NB: We've left the time-point as a character string, but we can fix this before plotting

```{r}
pcr_long %>%
  mutate(TimePoint = str_replace(TimePoint, "hr", "")) %>%
  ggplot(aes(x = as.integer(TimePoint), y =Ct, colour =Gene)) +
  geom_line() +
  geom_point() +
  facet_wrap(~CellType, ncol = 1) +
  labs(x = "Time (hr)")
```

