---
title: "R-Adelaide Workshop"
author: 'University of Adelaide, Bioinformatics Hub'
date: "July 20-22, 2016"
output: html_document
---

## Provisional Timetable

| **Time**   | **Wednesday**                          | **Thursday**                      | **Friday**                   |
| ------ | ---------------------------------- | ----------------------------- | ------------------------- |
|9:00am  | Introduction to `R Studio` <br> Data Import In R  | Understanding Data Types in `R` | `R` In The Genomics Era <br> (The Bioconductor Project) |  
|10:30am | *Morning Tea*                      | *Morning Tea*                   | *Morning Tea*                    |
|11:00am | Manipulating Data Using `dplyr` and `reshape2` | Writing Functions in `R`        | `R` In the Genomics Era <br> (The Bioconductor Project)  |
|12:30pm | *Lunch*                            | *Lunch*                         | *Lunch*                          |
|1:30pm  | Data Visualisation Using `ggplot2` <br> Making Publication Quality Images | Getting the Most Out Of Functions <br> Parallel Processing | Basic Statistical Techniques Using `R`             |
|3:00pm  | *Coffee Break*                     | *Coffee Break*                | *Coffee Break*                 |
|3:30pm  | Writing Reports Using `R Markdown` | Running `R Studio` In The Cloud | Mixed-Effects Modelling |

# Computer Setup

## R

Please ensure you have the latest version of `R`.
The current version (`R-3.3.0`) is available at https://www.cran.r-project.org/
Install this before installing `R Studio


## R Studio
Please ensure you have the latest version of `R Studio`
The current version(`RStudio 0.99.902`) is available at https://www.rstudio.com/

R Studio will usually automatically detect your most recent version of R and will use this by default.

## R Packages
To install all required packages, please execute the following commands in R:
```{r, eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite(c("dlpyr", "reshape2", "readr", "readxl", "ggplot2", "knitr", "parallel", "snow", "biomaRt", "GenomicRanges", "nlme", "lme4", "lattice", "MASS"))
```

If having problems with the installation, please try:
```{r, eval=FALSE}
biocValid(fix = TRUE)
```


# Notes for Windows Users

Installing the package `dplyr` can be problematic on University laptops running Windows.
If you are having problems, follow these steps:

1. Create a new folder `C:\R`  
2. Create a new `Environment Variable` (see below) called `R_LIBS_USER` and give it the value `C:\R`


## Windows 10 and Windows 8

1. In Search, search for and then select: `System (Control Panel)`
2. Click the `Advanced` system settings link.
3. Click `Environment Variables`. 
4. In the User Variables Section, click `New...`
5. Name the New Variable `R_LIBS_USER` and specify the value `C:\R`

## Windows 7

1. From the desktop, right click the Computer icon.
2. Choose Properties from the context menu.
3. Click the `Advanced` system settings link.
4. Click `Environment Variables`. 
5. In the User Variables Section, click `New...`
6. Name the New Variable `R_LIBS_USER` and specify the value `C:\R`

