---
title: "R-Adelaide Workshop"
author: 'University of Adelaide, Bioinformatics Hub'
date: "July 20-22, 2016"
output: html_document
---

# Computer Setup

## R

Please ensure you have the latest version of `R`.
The current version (`R-3.3.1`) is available at http://cran.ms.unimelb.edu.au/
Install this before installing `R Studio`


## R Studio
Please ensure you have the latest version of `R Studio`
The current version(`RStudio 0.99.902`) is available at [https://www.rstudio.com/](https://www.rstudio.com/products/rstudio/download/)

R Studio will usually automatically detect your most recent version of R and will use this by default.

## R Packages
To install all required packages, please execute the following commands in R (You can copy & paste both lines if you'd like):
```{r, eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite(c("dlpyr", "reshape2", "readr", "readxl", "ggplot2", "knitr", "parallel", "snow", "biomaRt", "GenomicRanges", "nlme", "lme4", "lattice", "MASS", "rmarkdown", "lmerTest", "car", "tibble"))
```

If you're having problems with the installation, please try:
```{r, eval=FALSE}
biocValid(fix = TRUE)
```


# Notes for Windows Users

Installing the package `dplyr` can be problematic on University laptops running Windows.
If you are having problems, follow these steps:

1. Create a new folder `C:\R`  
2. Create a new `Environment Variable` (see below) called `R_LIBS_USER` and give it the value `C:\R`.
The steps below can guide you through this.

## Windows 10 and Windows 8

1. In Search, search for and then select: `System (Control Panel)`
2. Click the `Advanced` system settings link.
3. Click `Environment Variables`. 
4. In the User Variables Section, click `New...`
5. Name the New Variable `R_LIBS_USER` and specify the value `C:\R`

## Windows 7

1. From the desktop, right click the Computer icon.
2. Choose Properties from the context menu.
3. Click the `Advanced` system settings link.
4. Click `Environment Variables`. 
5. In the User Variables Section, click `New...`
6. Name the New Variable `R_LIBS_USER` and specify the value `C:\R`

